### 2022-05-25 diagram_sub.tex
lots to commit

\documentclass[tikz,border=2mm]{standalone}

\usepackage{grffile}
\usetikzlibrary{arrows,positioning}

\begin{document}

\begin{tikzpicture}[node distance=3.5cm,auto,>=latex',every node/.append style={align=center},int/.style={draw, minimum size=1cm},inverter/.style={rectangle,draw,inner sep=2pt,minimum size=6mm},]
    \node [int, ultra thick] (S)             {$S$};
    \node [int, right=of S, ultra thick] (E)             {$E$};
    \node [int, right=of E, ultra thick] (Ip)             {$Ip$};
    \node [int, above right=of Ip, ultra thick] (Ia)             {$I_a$};
    \node [int, below right=of Ip, ultra thick] (Is)             {$I_s$};
    \node [int, right=of Ia, ultra thick] (R)             {$R$};
    \node [int, right=of Is, ultra thick] (D)             {$D$};
    \path[->, ultra thick] (S) edge node [above] {$\lambda(t)$} (E);
    \path[->, ultra thick] (E) edge node [above] {$\sigma$} (Ip);
    \path[->, ultra thick] (Ip) edge node [sloped, above] {$p \mu$} (Ia);
    \path[->, ultra thick] (Ip) edge node [sloped, above] {$(1-p) \mu$} (Is);
    \path[->, ultra thick] (Ia) edge node [above] {$\gamma_a$} (R);
    \path[->, ultra thick] (Is) edge node [sloped, above] {$(1-f) \gamma_s$} (R);
    \path[->, ultra thick] (Is) edge node [above] {$f \gamma_s$} (D);
    \node [right=1.5cm of S, ultra thick] (Ts)             {};
    \path[->, ultra thick, dashed] (Ia.west) edge [out=180, in=90] node [sloped, above] {$\beta_a$} ([yshift=5mm]Ts.north);
    \path[->, ultra thick, dashed] (Ip.north) edge [out=90, in=90] node [sloped, above] {$\beta_p$} ([yshift=5mm]Ts.north);
    \path[->, ultra thick, dashed] (Is.west) edge [out=180, in=-90] node [sloped, above] {$(1-\delta) \beta_s$} (Ts.south);
    \node [below=15cm of S, anchor=west] (pic) {\includegraphics{figure_sub.ggp.pdf}};
\end{tikzpicture}
\end{document}